---
title: "Open Stats Lab One-way ANOVA in R Using Data from Bastian, Jetten, & Ferris (2014)"
author: "Cory J. Cascalheira"
date: "2018-11-12"
slug: osl-bastian-jetten-ferris-2014
categories:
  - R
  - Tutorial
tags:
  - anova
  - t-test
image:
  caption: ""
  focal_point: ""
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Open Stats Lab
___
Kevin P. McIntyre developed this amazing resource for students of psychology. Check out [Open Stats Lab](https://sites.trinity.edu/osl/) for a collection of all activities.

Each activity includes an article from *Psychological Science*, a data set, and an activity to complete in SPSS. However, if you are an open source fanatic, you could also complete the activity in [JASP](https://jasp-stats.org/). For tips on how to use JASP, check out [this resource](https://osf.io/t56kg/) created by Buchanan, Hopke, and Donaldson (2018).

I prefer to get my hands deep into the data. Dr. McIntyre does not *yet* offer an R activity to accompany the work of [Bastian, Jetten, and Ferris (2014)](https://journals.sagepub.com/stoken/default+domain/yFi5kZDGWMGBNQY62zIE/full), so here is one possible solution written in R.

# Analysis 
___
I will perform assumption checks for each test prior to running it. We already know that the data meet all assumptions, otherwise the authors would have used a different analytic approach. However, checking the assumptions is helpful because:

1. reproducibility and accuracy can be verified; and
2. if you are a student, then you should form the habit of testing assumptions.

This analysis will follow the data science workflow advocated by [Garrett Grolemund and Hadley Wickham](https://r4ds.had.co.nz/introduction.html). First, we will set-up our session and import the data. Then, we must clean the data. Next, we will transform, model, and visualize the data to understand it. Finally, we will communicate our findings.

Our task is to perform four analyses.

1. **Independent t-Test** - did participants differ significantly in their experience of pain intensity and pain unpleasantness?
2. **Independent t-Test** - did participants experience significant differences in affect?
3. **Independent t-Test** - did participants in the pain condition find their activity more threatening and challenging?
4. **One-way ANOVA** - does pain lead to differences in group bonding?

## Import
___
Let's load the packages necessary for this analysis into our workspace.

```{r message = FALSE}
library(tidyverse)
library(psych)
library(car)
```
