---
date: "2020-12-11"
draft: false
linktitle: Fall 2020
menu:
  teaching:
    parent: Course Evaluations
    weight: 1
title: Fall 2020
type: docs
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, message=FALSE}
# Load dependencies 
library(tidyverse)
library(lubridate)
library(knitr)
library(kableExtra)

# Import data
survey <- read_csv("data/NMSU_CEP-320_Fall-2020_Evaluations.csv")

# Drop survey identifier & student ID & complete
survey_1 <- survey[,-c(2, 4, 11)]

# Clean names vector
survey_names <- c("record", "time", "organized_content", "concepts_clear", "helpful", "feedback", "overall", "improve")

# Clean the names
names(survey_1) <- survey_names

# Transfer date-time to month
survey_2 <- survey_1 %>%
  mutate(time = month(time)) %>%
  # Recode scale anchors to numeric
  mutate(
    organized_content = recode(organized_content, "Strongly Disagree" = 1, "Disagree" = 2, "Somewhat Disagree" = 3, "Neutral" = 4, "Somewhat Agree" = 5, "Agree" = 6, "Strongly Agree" = 7),
    concepts_clear = recode(concepts_clear, "Strongly Disagree" = 1, "Disagree" = 2, "Somewhat Disagree" = 3, "Neutral" = 4, "Somewhat Agree" = 5, "Agree" = 6, "Strongly Agree" = 7),
    helpful = recode(helpful, "Strongly Disagree" = 1, "Disagree" = 2, "Somewhat Disagree" = 3, "Neutral" = 4, "Somewhat Agree" = 5, "Agree" = 6, "Strongly Agree" = 7),
    feedback = recode(feedback, "Strongly Disagree" = 1, "Disagree" = 2, "Somewhat Disagree" = 3, "Neutral" = 4, "Somewhat Agree" = 5, "Agree" = 6, "Strongly Agree" = 7),
    overall = recode(overall, "Very Poor" = 1, "Poor" = 2, "Questionable" = 3, "Average" = 4, "Good" = 5, "Great" = 6, "Excellent" = 7)
  )
```

## CEP 320: Sex Roles in Education
___

This semester, I taught **CEP 320: Sex Roles in Education** for the Department of Counseling & Educational Psychology in the College of Education at New Mexico State University. 

It was my first course as the instructor of record. 

Notable achievements included:

1. the improvement of the [syllabus](docs/CEP_320/CEP-320_Syllabus_Cascalheira.pdf), from an [older version](docs/CEP_320/CEP-320_Syllabus_Old.pdf);
2. the establishment of assignments linked to evidence-based [learning objectives](docs/CEP_320/CEP-320_Learning-Objectives.pdf);
3. the creation of a highly organized, asynchronous, [virtual course](https://www.youtube.com/watch?v=oh6ofPbOkrQ&list=PLLBryyM0S9nrCfe-JZpo_mYY0MniF5SFq) (Canvas IMSCC export file available unpon request); and
4. weekly announcements, [typically including video updates](https://www.youtube.com/watch?v=eGvEl4Nzlf8&list=PLLBryyM0S9nqtvCLj8hxvogYEbeWqPMmY).

### The Feedback
___

```{r echo=FALSE, message=FALSE}
# Percent participating
survey_percents <- survey_2 %>%
  mutate(time = recode(time, `12` = 11)) %>%
  count(time) %>%
  mutate(percent = n / 40)
```

Forty students were enrolled, which meant data for the evaluations were based on response rates of:

* mid-semester = `r survey_percents$percent[1] * 100`%
* end of semester = `r survey_percents$percent[2] * 100`%

#### Mid-Semester

As shown in the following table (which reports the mean; range = 1 to 7), students tended to agree strongly with all criteria of instructor effectiveness. On average, they rated my overall effectiveness as excellent.

```{r echo=FALSE, message=FALSE}
# Prepare data frame
survey_2[, -c(1, 8)] %>%
  filter(time == 10) %>%
  select(-time) %>%
  gather(key = "criterion", value = "score") %>%
  group_by(criterion) %>%
  summarize(Average = mean(score, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(criterion = recode(criterion, "concepts_clear" = "Concepts Clear", "feedback" = "Constructive Feedback", "helpful" = "Timely Help", "organized_content" = "Organized Content", "overall" = "Overall"),
         Average = format(round(Average, 2))) %>%
  rename("Criterion" = criterion) %>%
  kable() %>%
  kable_styling(c("striped", "hover"))
```

The following qualitative feedback came from the open-ended responses on the survey.

**STRENGTHS**

"I feel the instructor has done great with communication and feedback from journals. He responds in a timely manner to emails and is always there to help if we have questions or need assistance. There is not anything they could improve on in my view."

"The weekly announcements really help!"

"He was absolutely amazing! Even though this class was online I really felt that I learned quite a bit through the class. 10/10 Work queen. Also I think that the feedback that he offered on every assignment was very informative and helped me improve on my work for the next time."

"This is my favorite professor I have ever had. I think the feedback that he had was phenomenal and they really make my day. I highly enjoy the assignments we have and how open he is in expressing his thoughts in comparison to ours. 15/10 would recommend."

"I think the instructor does well in the class, through reaching out to us every week and giving us reminders about due dates. He also offers assistance if we have questions and replies in a timely manner. I enjoy the class and the instructors involvement even though class is asynchronous online."

"I absolutely love this course." 

"I feel like the instructor has done a great job teaching this course online, therefore, I don't necessarily think he should do anything in improving the course material. I'm learning and attaining the information very well by the way he is delivering the content for this course."

"I don't think there needs to be any improvements! This course was very clear and helpful!"

"Cory is doing a wonderful job so far. I enjoy class assignments and with each assignment my worldview has grown and increased."

"Doing the weekly announcements really helps. Doing a video for the bigger projects helps too." 

"Nothing, this class is easy to navigate, organized and straight forward. It is actually the best class that I have come across in terms of structure."

"Professor Corey sets up this course so well, I am never confused. This is my favorite course I have ever taken."

"This instructor has been great and very helpful through out the course."

**CONCERNS & AREAS OF GROWTH**

"I feel like it's easier to have the discussion and reply to the discussion due on the same day. It's not a big deal at all, but it could easily be done all at once."

"Nothing, perhaps doing video lectures."

"Could improve in giving better feedback, especially to discussions or reaction journals :)"

"I think Cory does a great job. If anything, i would suggest lecturing; however, that is just a personal preference."

"Possibly uploading lectures instead of having students completely read from the book for their own learning." 

"The whole class is structure very organized matter. The one thing I would suggested in improving is making the class more 'fun'. maybe adding more visuals."

"Allow breaks from regular work when big projects are due that week"

"Maybe a little more time with the instructor via zoom."

#### End of Semester

By the end of the semester, the ratings increased, but fewer students completed the survey.

```{r echo=FALSE, message=FALSE}
# Prepare data frame
survey_2[, -c(1, 8)] %>%
  filter(time > 10) %>%
  select(-time) %>%
  gather(key = "criterion", value = "score") %>%
  group_by(criterion) %>%
  summarize(Average = mean(score, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(criterion = recode(criterion, "concepts_clear" = "Concepts Clear", "feedback" = "Constructive Feedback", "helpful" = "Timely Help", "organized_content" = "Organized Content", "overall" = "Overall"),
         Average = format(round(Average, 2))) %>%
  rename("Criterion" = criterion) %>%
  kable() %>%
  kable_styling(c("striped", "hover"))
```

The following qualitative feedback came from the official course evaluations from the department.